Individual Variation in Animal Movement and Effects on Seed Dispersal Distance

### 2021-01-08 paper/paper.Rmd
update to model processes

To take advantage of all the locations recorded, despite this variation in duration of time intervals, we calculated rates of movement, as the average number of meters moved per minute over the entire tracking period for that individual bird. These movement rates were later used in our simulation models as the parameter to describe the probability distribution of movement distances at each one minute time step for each simulation run. In the case of our null model, the movement rate was averaged across all individuals for all the tracking periods, and at the social group level we took the average number of meters moved per minute over the tracking periods for all individuals belonging to that group, with a total of 7 independent groups.

Parameter values for gut retention time were also based on empirically collected data from previous studies [@holbrook_using_2007], were  passage trials were carried out with *Pteroglossus pluricinctus* individuals. However, these passage trials were not performed on every individual captured and some of the trials were also performed with captive birds, thus we could not explore individual variation in gut retention times for the individuals in the movement data set. These trials showed that the average gut retention time for *Pterglossus pluricinctus* was 28 minutes [@holbrook_using_2007], and maximum retention times over 100 minutes. The distribution of gut retention times is characteristically fat-tailed, thus we used a gamma distribution (shape = 4, scale = 5) with the appropriate shift to match our average retention time of 28 minutes [@morales_effects_2006; @levey_effects_2005].

@@

We give randomly between 3-5 seeds, because that's what's observed in the field. We are simulating here the trajectory and fate of seeds after only one event of frugivory to make things simpler.

We have a maximum landscape area, based on the home range size of the individuals, we chose a grid, and it's modled as a torus so that there is no edge effect.? But do I want this if we want to explore long distance?

>>>>

To take advantage of all the locations recorded for series of successive points, despite this variation in duration of time intervals, we calculated rates of movement, as the average number of meters moved per minute over the entire tracking period for that individual bird. These movement rates were later used in our simulation models as the parameter to describe the probability distribution of movement distances at each one minute time step for each simulation run. In the case of our null model, the movement rate was averaged across all individuals for all the tracking periods, and at the social group level we took the average number of meters moved per minute over the tracking periods for all individuals belonging to that group, with a total of 7 independent groups.

Parameter values for gut retention time were also based on empirically collected data from previous studies [@holbrook_using_2007], were  passage trials were carried out with *Pteroglossus pluricinctus* individuals. However, these passage trials were not performed on every individual captured and some of the trials were also performed with captive birds, thus we could not explore individual variation in gut retention times for the individuals in the movement data set. These trials showed that the average gut retention time for *Pterglossus pluricinctus* was 28 minutes [@holbrook_using_2007], and maximum retention times over 100 minutes. The distribution of gut retention times is characteristically fat-tailed, thus we used a gamma distribution (shape = 4, scale = 5) with the appropriate shift to match our average retention time of 28 minutes [@morales_effects_2006; @levey_effects_2005], and used this distribution to sample gut retention times for each of our simulation runs.

@@

Each simulation run in our models consisted of one animal in a homogeneous landscape starting at the focus tree located at coordinates (0,0). We assume the landscape is in meters, as the sampling of movement distances for animal movement comes from estimates of meters moved per minute, the movement rates.  Informed from field observations, *Pteroglossus pluricinctus* have a mean visit length of 4.0 minutes for each fruiting tree, consuming between 2 and 5 seeds during each visit, and not visiting another fruiting tree immediately after feeding (Holbrook's dissertation and personal communication [@holbrook2007seed]). Due to this underlying behavior, we decided to focus our simulations specifically to the events associated to feeding from one focus tree, and the animal movement occurring after foraging and before visiting another fruiting tree. Since one of the objectives for this paper is to explore the occurrence of long-distance dispersal events, we did not set boundaries for the landscape. *Pteroglossus pluricinctus* has shown maximum travel distances exceeding 3500m in a single 30 minute tracking interval, thus showing the potential to disperse *Virola* seeds at long ranges [@holbrook_home_2011].

At the beginning of each simulation run, the animal consumed five seeds at the focus tree location, and each of these seeds was assigned a gut retention time sampled from a gamma distribution (shape=4, scale=5, shift=8). Once the animal in the simulation consumed the seeds, a movement distance was sampled from the assigned exponential distribution (which varied depending on the underlying model such as the null, individual or family group model) and a random direction from $0^o$ to $360^o$ degrees was used to determine the animal's position for the next time step, repeating this process for every time step and thus following an uncorrelated random walk with no directional tendency. Once the simulation run's time matched the gut retention time for a seed, that seed would get dropped at the animal's location at that specific time point, thus allowing us to record seed location in the landscape. For each of our models we ran 10,000 simulation runs per individual or family group, depending on the model, and collected information on animal and seed locations at every time step.

## Seed dispersal and aggregation metrics

Bring up that each seed gets dropped when the simulation reaches the seed's previously determined gut retention time, and thus the seed's location is the lcoation of the bird at the simulation time that matches the gut retention time. The simulation run ends when the last seed is dropped, which is determined by the largest gut retention time. 
We did not have limits set on the landscape, as we are trying to also understand the effects of this variation on the potential for long-distance dispersal events in toucan-generated seed dispersal curves. Maximum recorded distances for these toucans in the datasets is of 3665m in a single 30 minute tracking interval, so we didn't want to limit the dispersal potential by setting the landscape to a torus and constricting its size.
